---
title: "Math 490 HW #2"
author: "Maxwell Levin"
date: "January 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(comment=NA)
```

## Question 1.

A box contains 7 balls numbered 1 through 7. Alice chooses two balls at random with replacement, and label the two observed values X and Y. Let V = X - 2Y and let W = |X - Y|.



####a. Find the p.m.f. of V.

```{r}
table <- matrix(c("1/49", "1/49", "2/49", "2/49", "3/49", "3/49", "4/49", "3/49", "4/49", "3/49", "4/49", "3/49", "4/49", "3/49", "3/49", "2/49", "2/49", "1/49", "1/49"), ncol=19, byrow=TRUE)
colnames(table) <- c("-13", "-12", "-11", "-10", "-9", "-8", "-7", "-6", "-5", "-4", "-3", "-2", "-1", "0", "1", "2", "3", "4", "5")
rownames(table) <- c("p.m.f.")
table <- as.table(table)
table
```





####b. Find E(V) and Var(V).

Since X & Y are independent events we can write:
  
  \[
  E(X - 2Y) = E(X) - 2E(Y),
  \]
  \[
  Var(X - 2Y) = Var(X) + 4Var(Y).
  \]
  
Furthermore, since E(X) = E(Y) and Var(X) = Var(Y) we have:
  \[ 
  E(V) = -E(X),
  \]
  \[
  Var(V) = 5Var(X).
  \]
Thus 

\[
E(V) = -\frac{1 + 2 + \dots + 7}{7} = -4,
\]
  
and 
\[
Var(V) = 5\frac{(1 - 4)^2 + (2 - 4)^2 + \dots + (7 - 4)^2]}{7} = 20.
\]


####c. Use R to sample 1024 values from V, make a histogram, and report the mean and variance of the observed values.

```{r, echo=TRUE}
X <- sample(1:7, 1024, replace=T)
Y <- sample(1:7, 1024, replace=T)
V = X - 2*Y
hist(V, breaks=c(-13.5, -12.5, -11.5, -10.5, -9.5, -8.5, -7.5, -6.5, -5.5, -4.5, -3.5, -2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5), prob=F)
```

The mean is:
```{r}
mean(V)
```
The variance is:
```{r}
var(V)
```


####d. Find the p.m.f. of W.

```{r}
wTab <- matrix(c("7/49", "12/49", "10/49", "8/49", "6/49", "4/49", "2/49"), ncol=7, byrow=T) 
colnames(wTab) <- c("0", "1", "2", "3", "4", "5", "6")
rownames(wTab) <- c("p.m.f.")
wTab <- as.table(wTab)
wTab
```



####e. Find E(W) and Var(W).

We can compute E(W) using our table above:

\[
  E(W) = 0\left(\frac{7}{49}\right) + 1\left(\frac{12}{49}\right) + \dots + 6\left(\frac{2}{49}\right).
\]

This yields:
\[
  E(W) \approx 2.286.
\]

We can compute Var(W) using our table and E(W):
\[
  Var(W) = (0 - 2.286)^2\left(\frac{7}{49}\right) + (1 - 2.286)^2\left(\frac{12}{49}\right) + \dots + (6 - 2.286)^2\left(\frac{2}{49}\right).
\]

This yields:
\[
Var(W) \approx 2.776.
\]

####f. Use R to sample 1024 values from W, make a histogram, and report the mean and variance of the observed values.

```{r, echo=TRUE}
X <- sample(1:7, 1024, replace=T)
Y <- sample(1:7, 1024, replace=T)
W = abs(X - Y)
hist(W, breaks=c(-0.5, 0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5), prob=F)
```

The mean is:
```{r}
mean(W)
```
The variance is:
```{r}
var(W)
```




