---
title: 'Math 490 HW #4'
author: "Maxwell Levin"
date: "January 31, 2018"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/', warning=FALSE, message=FALSE)
knitr::opts_chunk$set(comment=NA)
```

## Question 1.

####Let random variable U be uniformly distributed over the interval [a, b]. Find $\small{E(U^2)}$ and $\small{Var(U^2)}$.

Since $\small{U}$ is a continuous random uniform distribution, we calculate $\small{E(U^2)}$ using the equation

\[
  E(U^2) = \int_a^bU^2f(U) dU,
\]
\[
  E(U^2) = \frac{1}{b-a}\int_a^b U^2 dU,
\]
\[
  E(U^2) = \frac{a^2 + ab + b^2}{3}.
\]


We now calculate the variance of $\small{U^2}$:

\[
  Var(U^2) = E(U^4) - E(U^2),
\]
\[
  Var(U^2) = \frac{b^5 - a^5}{5(b-a)} - \left(\frac{b^3-a^3}{3(b-a)}\right)^2.
\]

## Question 2.

####Consider sampling n = 4 random numbers $\small{U_1, U_2, U_3, U_4}$ from the Uniform[0,1] distribution, and let

\[
  \overline{V_4} = \frac{U_1^2 + U_2^2 + U_3^2 + U_4^2}{4}.
\]

#### a. What are the mean and standard deviation of $\small\overline{V_4}$?

We calculate the mean:

\[
  E(\overline{V_4}) = E\left(\frac{U_1^2 + U_2^2 + U_3^2 + U_4^2}{4}\right),
\]
\[
  E(\overline{V_4}) = E\left(\frac{1}{4}E(U_1^2) + \frac{1}{4}E(U_2^2) + \frac{1}{4}E(U_3^2) + \frac{1}{4}E(U_4^2)\right),
\]
\[
  E(\overline{V_4}) = E(U^2).
\]

Thus we can use our result from question 1 with $\small{a = 0, b = 1}$ to get

\[
E(\overline{V_4}`b) = \frac{1}{3}.
\]

To calculate variance we use

\[
  Var(\overline{V_4}) = Var\left(\frac{U_1^2 + U_2^2 + U_3^2 + U_4^2}{4}\right),
\]
\[
  Var(\overline{V_4}) = \left(\frac{1}{4}\right)^2Var(U_1^2) + \left(\frac{1}{4}\right)^2Var(U_2^2) + \left(\frac{1}{4}\right)^2Var(U_3^2) + \left(\frac{1}{4}\right)^2Var(U_4^2),
\]
\[
  Var(\overline{V_4}) = \frac{1}{4}Var(U^2).
\]

Thus we can again apply our results from question 1 to achieve

\[
  Var(\overline{V_4}) = \frac{1}{45} \approx 0.0222,
\]
\[
  SD(\overline{V_4}) = \sqrt{Var(\overline{V_4})} \approx 0.1491
\]


#### b. Use R to simulate 1024 sample means $\small\overline{V_4}$?, make a histogram, and report the mean (center) and standard deviation (spread) of the simulated sample means.

```{r}
u1 = runif(1024, 0, 1)
u2 = runif(1024, 0, 1)
u3 = runif(1024, 0, 1)
u4 = runif(1024, 0, 1)

Vbar = (u1^2 + u2^2 + u3^2 + u4^2) / 4

hist(Vbar)


```
The Mean is
```{r, echo=FALSE}
mean(Vbar)
```


The Standard Deviation is
```{r, echo=FALSE}
SD = sqrt(sum((Vbar - mean(Vbar))^2)/1024)
SD

```





#### c. Use a normal probability plot to access the normality of the 1024 sample means simulated in part b.

```{r, echo=FALSE}
qqnorm(Vbar, col="orange")
qqline(Vbar, col="blue")
```

Our normal probability plot indicates that our sample means are pretty close to a normal distribution.


#### d. Use the Shapiro-Wilk test to assess the normality of the 1024 sample means simulated in part b.

```{r, echo=FALSE}
shapiro.test(Vbar)
```

Our Shapiro-Wilk test tells us that there is very strong evidence for normality of our sample means.

## Question 3.

####Consider sampling n = 25 random numbers $\small{U_1, U_2, \dots, U_25}$ from the Uniform[0,1] distribution, and let

\[
  \overline{V_{25}} = \frac{U_1^2 + U_2^2 + \dots + U_{25}^2}{25}.
\]

#### a. What are the mean and standard deviation of $\small\overline{V_4}$?

We calculate the mean:

\[
  E(\overline{V_{25}}) = E\left(\frac{U_1^2 + U_2^2 + \dots + U_{25}^2}{25}\right),
\]
\[
  E(\overline{V_{25}}) = E\left(\frac{1}{25}E(U_1^2) + \frac{1}{25}E(U_2^2) + \dots + \frac{1}{25}E(U_{25}^2)\right),
\]
\[
  E(\overline{V_{25}}) = E(U^2).
\]

Thus for an interval from 0 to 1 our mean is

\[
  E(\overline{V_{25}}) = \frac{1}{3}.
\]

We calculate the variance:

\[
  Var(\overline{V_{25}}) = Var\left(\frac{U_1^2 + U_2^2 + \dots + U_{25}^2}{25}\right)
\]
\[
  Var(\overline{V_{25}}) = \left(\frac{1}{25}\right)^2Var(U_1^2) + \left(\frac{1}{25}\right)^2Var(U_2^2) + \dots +
  \left(\frac{1}{25}\right)^2Var(U_{25}^2)
\]
\[
  Var(\overline{V_{25}}) = \frac{1}{25}Var(U^2).
\]

We can apply our results from question 2 to get

\[
  Var(\overline{V_{25}}) = \frac{4}{1125} \approx 0.0036.
\]

That is, the standard deviation of our sample means when n = 25 is

\[
  SD(\overline{V_{25}}) = \sqrt{Var(\overline{V_{25}})} \approx 0.0596.
\]


#### b. Use R to simulate 1024 sample means $\small\overline{V_{25}}$?, make a histogram, and report the mean (center) and standard deviation (spread) of the simulated sample means.


```{r}
clt = function(sam, rep) {
  obs = NULL
  for (i in 1:rep) {
    v = runif(sam, 0, 1)^2
    Vbar = mean(v)
    obs = c(obs, Vbar)
  }
  obs;
}

simul = clt(25, 1024)

hist(simul)
```

The mean is
```{r, echo=FALSE}
mean(simul)
```

The standard deviation is
```{r, echo=FALSE}
SD = sqrt(sum((simul - mean(simul))^2)/1024)
SD
```


#### c. Use a normal probability plot to access the normality of the 1024 sample means simulated in part b.

```{r,echo=FALSE}
qqnorm(simul, col="orange")
qqline(simul, col="blue")
```

Our normal probability plot suggests that our simulation of sample means is pretty close to a normal distribution.

#### d. Use the Shapiro-Wilk test to assess the normality of the 1024 sample means simulated in part b.

```{r, echo=FALSE}
shapiro.test(simul)
```

Our Shapiro-Wilk normality test gives us no evidence for normality since the P-value is larger than 0.1.


