---
title: 'Math 490 HW #17'
author: "Maxwell Levin"
date: "April 3, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, fig.width = 6, fig.height = 4, fig.align = "center")
```

## Question 1.

#### Use our class data to carry out the permuation test on whether gender and sleep are independent. Report the P-value and draw your conclusion at a level of significance of $\small{\alpha = 10\%}$.

First we state our hypotheses; our null hypothesis is that gender and sleep are independent and our alternative hypothesis is that gender and sleep are not independent.

Now we can run a permutation test on whether gender and sleep are independent by running the following code in R:
```{r, echo=TRUE}
# making a 2 x 2 table
original = matrix(c(4, 5, 8, 5), nrow = 2, byrow = T)  

# computing the likelihood of the observed table
obsProb = choose(9, 4) * choose(13, 8) / choose(22, 12) 

permTest = function(x, y, rep) {  
  # x is the category label like treatments
  # y is the category variable like outcomes
	count = 0
	for (i in 1:rep) {
		xPerm = sample(x)  # a random permuation of label x
		dataPerm = data.frame(y, xPerm)
		tab = table(y, xPerm)
		# print(tab)   # use this to verify the permuted table
		if (choose(sum(tab[1,]), tab[1,1]) * choose(sum(tab[2,]), tab[2,1]) 
		    / choose(sum(tab), sum(tab[,1])) <= obsProb) {
		  count = count + 1
		}
	}
	count / rep
}

# Get the data
our_data = read.table("math490.R", header=TRUE)
attach(our_data)

outcome = rep(c("Early", "Night"), c(9, 13))
treatment = rep(c("Female", "Male"), c(12, 10))
outAndTreat = data.frame(gender, sleep_type)
permTest(gender, sleep_type, 1024)
```

This gives us a P-value of about 0.6, which is way larger than our $\small{\alpha}$ value. Thus we do not have sufficient evidence to reject the null hypothesis that gender and sleep are independent.

