---
title: 'Math 490 HW #16'
author: "Maxwell Levin"
date: "March 30, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, fig.width = 6, fig.height = 4, fig.align = "center")
```


## Question 1. 
#### Use our class data to carry out the chi-square test on whether gender and sleep are independent. Follow the 4-step format.

#### Step 1: State the null and alternative hypotheses
Since we are asked to test whether gender and sleep are independent attributes, a good place to start for our null hypothesis is that gender and sleep are not independent. This means that our alternative hypothesis is that gender and sleep are independent attributes.


#### Step 2: Summarize the data using a contingency table

This problem is simple enough that we do not necessarily need R to generate our contingency table, but nonetheless I have written the following code in R as good practice and to save time in the future should I need to create a larger contingency table later on.



```{r, echo=TRUE}
# Import our class data
our_data = read.table("math490.R", header=TRUE)
attach(our_data)

# Splice the data for our contingency table
early_m = our_data[ which(sleep_type == 'early' & gender == 'm'), ]
early_f = our_data[ which(sleep_type == 'early' & gender == 'f'), ]
night_m = our_data[ which(sleep_type == 'night' & gender == 'm'), ]
night_f = our_data[ which(sleep_type == 'night' & gender == 'f'), ]

# Calculate the totals (num_early + num_night and num_m + num_f should both be 22)
num_early = nrow(early_m) + nrow(early_f)
num_night = nrow(night_m) + nrow(night_f)
num_m = nrow(our_data[ which(gender == 'm'), ])
num_f = nrow(our_data[ which(gender == 'f'), ])

obs = c(nrow(early_m), nrow(early_f), nrow(night_m), nrow(night_f)) # I'll use this later

# Create the table
makeTable = function(value, totals) { 
  # Must have exactly 1 degree of freedom
  # Value is the top left value in the table
  # Totals are the sum of row elements in decending order, then the columns in ascending order.
  v2 = totals[1] - value
  v3 = totals[3] - value
  v4 = totals[2] - v3
  ans = matrix(c(value, v2, totals[1], 
                 v3, v4, totals[2], totals[3], 
                 totals[4], totals[1] + totals[2]), 
               ncol=3)
  colnames(ans) = c("Male", "Female", "Total")
  rownames(ans) = c("Early", "Night", "Total")
  ans
}
totals = c(num_early, num_night, num_m, num_f)
as.table(makeTable(5, totals))
```


Now we run the following code in R to compute the expected frequencies
```{r, echo=TRUE}
total = num_m + num_f
row1 = c(num_early * num_m / total, num_early * num_f / total)
row2 = c(num_night * num_m / total, num_night * num_f / total)
exp = c(row1, row2) # I'll use this later
exp_table = matrix(c(row1, row2), ncol=2)
colnames(exp_table) = c("Male", "Female")
rownames(exp_table) = c("Early", "Night")
as.table(exp_table)
```



#### Step 3: Compute the chi-square statistic 

We know from class that we can compute this by

\[
\chi^2 = \sum\limits_{i=1}^r \sum\limits_{j=1}^k \frac{(Obs_{i,j} - Exp_{i,j})^2}{Exp_{i,j}}.
\]

We run the following code in R to compute our chi-square statistic with 1 degree of freedom
```{r, echo=TRUE}
chi_stat = sum(((obs - exp)^2) / exp)
chi_stat
```

#### Step 4: Assess the P-value from our chi-square statistic

We can compute our P-value by running the following code in R
```{r, echo=TRUE}
1 - pchisq(chi_stat, 1)
```
We see that this is much bigger than a 5% significance level. Thus we do not have enough evidence to reject the null hypothesis that sleep and gender are not independent. 


## Question 2. 
#### Use our class data to carry out Fisher's exact test on whether gender and sleep are independent. List all possible contingency tables and compute the P-value.




To do this we must list all the possible contingency tables. We can do this in R quite easily:
```{r, echo=TRUE}
makeTable = function(value, totals) { 
  # Must have exactly 1 degree of freedom
  # Value is the top left value in the table
  # Totals are the sum of row elements in decending order, then the columns in ascending order.
  v2 = totals[1] - value
  v3 = totals[3] - value
  v4 = totals[2] - v3
  ans = matrix(c(value, v2, v3, v4), ncol=2)
  colnames(ans) = c("Male", "Female")
  rownames(ans) = c("Early", "Night")
  ans
}

fish = numeric(11)
for (i in 1:11) {
  matAns = makeTable(i-1, totals)
  fish[i] = choose(10, i-1) * choose(12, matAns[4]) / choose(22, 10)
  print(as.table(matAns))
  writeLines('This table has a probibility of:')
  print(fish[[i]])
  writeLines('')
}
```


We now must compute the sum of all the P-values (probabilities) less than or equal to the one we observed. We do this with the following command:
```{r, echo=TRUE}
sum(fish[fish <= 0.3086449])
```


Thus we see that our P-value is pretty close to what we calculated in the first question. Thus both the chi-square test and Fisher's exact test tell us that we do not have sufficient evidence to reject the null hypothesis. 




