---
title: "Math 490 HW #19"
author: "Maxwell Levin"
date: "April 9, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA, fig.width = 6, fig.height = 4, fig.align = "center")
```


## Question 1.

#### Use the method of solving a linear system to find the stationary distribution of the transition matrix in exercise 2.3 in Voss. Express your answer as a vector of fractions.

Voss provides us with the following transition matrix: 
\[
  \textbf{P} =
  \begin{bmatrix}
    \frac{2}{3} & \frac{1}{3} & 0 & 0 \\[1mm]
    \frac{1}{10} & \frac{9}{10} & 0 & 0 \\[1mm]
    \frac{1}{10} & 0 & \frac{9}{10} & 0 \\[1mm]
    \frac{1}{10} & 0 & 0 & \frac{9}{10} \\[1mm]
  \end{bmatrix}
\]

We can solve the linear system by taking the matrix product $\small{\bf{vP}}$ and setting it equal to $\small{\bf{v}}$, where $\small{\bf{v}}$ is the stationary distribution for the transition matrix $\small{\bf{P}}$. We do this manually first:

\[
  [a, b, c, d] 
  \begin{bmatrix}
    \frac{2}{3} & \frac{1}{3} & 0 & 0 \\[1mm]
    \frac{1}{10} & \frac{9}{10} & 0 & 0 \\[1mm]
    \frac{1}{10} & 0 & \frac{9}{10} & 0 \\[1mm]
    \frac{1}{10} & 0 & 0 & \frac{9}{10} \\[1mm]
  \end{bmatrix}
  = \left[\frac{2a}{3} + \frac{b}{10} + \frac{c}{10} + \frac{d}{10}, 
          \frac{a}{3} + \frac{9b}{10}, \frac{9c}{10}, \frac{9d}{10} \right].
\]

Since $\small{\textbf{v}}$ is a stationary vector for $\small{\bf{P}}$, the resulting product should be equal to $\small{\bf{v}}$. Thus we have
\[
  a = \frac{2a}{3} + \frac{b}{10} + \frac{c}{10} + \frac{d}{10},
\]
\[
  b = \frac{a}{3} + \frac{9b}{10},
\]
\[
  c = \frac{9c}{10},
\]
\[
  d =  \frac{9d}{10},
\]
\[
  a + b + c + d = 1.
\]

We see that both $\small{c}$ and $\small{d}$ must be 0, so this leaves
\[
  a = \frac{2a}{3} + \frac{b}{10},
\]
\[
  b = \frac{a}{3} + \frac{9b}{10},
\]
\[
  a + b = 1.
\]

This means that $\small{ a = \frac{3}{13}}$ and $\small{b = \frac{10}{13}}$. Thus our stationary distribution is

\[
  \textbf{v} = \left[ \frac{3}{13}, \frac{10}{13}, 0, 0  \right].
\]


## Question 2.

#### Use R to find the stationary distribution of the transition matrix in exercise 2.3. 

We run the following code in R:
```{r, echo=TRUE}
# Transition Matrix
P = matrix( c(2/3, 1/3, 0, 0, 1/10, 9/10, 0, 0, 1/10, 0, 9/10, 0, 1/10, 0, 0, 9/10),
            nrow=4, ncol=4, byrow=TRUE)

# Function that will compute the stationary vector for any
# row-oriented transition matrix
getEquilibrium = function(transition) {
  eigen_stuff = eigen(t(transition))
  eigen_vec1 = eigen_stuff$vectors[ , 1]
  library(MASS)
  fractions(abs(eigen_vec1) / sum(abs(eigen_vec1)))
}

getEquilibrium(P)
```


Thus we see that we've found the same stationary distribution as we found in the first question! This is good news.







