---
title: 'Math 490 HW #12'
author: "Maxwell Levin"
date: "March 5, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, fig.width = 6, fig.height = 6, fig.align = "center")
```

## Question 1.

#### Use our class data to construct:

#### a. 95% confidence interval for the mean heart rate for the female group

Recall from class that we found a 95% confidence interval to be given by,

\[
  \overline{Y}_n = \pm 1.96\frac{\sigma}{\sqrt{n}},
\]

where we approximate $\small{\sigma}$ by 
\[
  \sigma \approx S,
\]

where $\small{S}$ is our sample standard deviation. Thus our confidence interval is given by: 

\[
  \overline{Y}_n = \pm 1.96\frac{S}{\sqrt{n}}.
\]


We can grab our class data using R with the following code:

```
```{r}
our_data = read.table("math490.R", header=TRUE)
attach(our_data)
```


We can get a 95% confidence interval in R by simply asking R to run a t-test on our heartbeat data for females:
```{r}
f_hb = hb[gender == "f"]
t.test(f_hb)
```

Thus we see that a 95% confidence interval for the female group is the interval from 67.05 to 79.45. Additionally we also note that the mean heartbeat for females in Math 490 is 73.25 bpm.

#### b. 95% confidence interval for the mean heart rate for the male group

To do the same for the male group all we have to do is run the following code:
```{r}
m_hb = hb[gender == "m"]
t.test(m_hb)
```


We see that the interval from 55.77 to 72.42 is a 95% confidence interval for the mean heartrate of the male group. We also note that the mean heartrate for the male group is 64.1 bpm.


## Question 2.

#### Let $\small{X \stackrel{d}{\sim} Normal(0, 1)}$. Use Monte Carlo estimation to obtain an estimate for $\small{E(cos(X))}$ to three digits of accuracy.

























