---
title: 'Math 490 HW #18'
author: "Maxwell Levin"
date: "April 6, 2018"
header-includes:
  - \usepackage{amsmath}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, fig.width = 6, fig.height = 4, fig.align = "center")
```


## Question 1.

#### Use the transition matrix

\[
  P=
  \begin{bmatrix}
    \frac{2}{3} & \frac{1}{3} & 0 & 0 \\[1mm]
    \frac{1}{10} & \frac{9}{10} & 0 & 0 \\[1mm]
    \frac{1}{10} & 0 & \frac{9}{10} & 0 \\[1mm]
    \frac{1}{10} & 0 & 0 & \frac{9}{10} \\[1mm]
  \end{bmatrix}
\]

#### and the initial distribution, $\small{\mu = \left[ \frac{1}{4}, \frac{1}{4}, \frac{1}{4}, \frac{1}{4}\right]}$, to find the 10th-step transition matrix and the 15th-step distribution.

We can get the 10th-step transition matrix by running the following code in R:
```{r, echo=FALSE}
# install.packages("expm")
require("expm")
```


```{r}
# Transition Matrix
P = matrix( c(2/3, 1/3, 0, 0, 1/10, 9/10, 0, 0, 1/10, 0, 9/10, 0, 1/10, 0, 0, 9/10),
            nrow=4, ncol=4, byrow=TRUE)

# Initial distribution
mu = c(1/4, 1/4, 1/4, 1/4)

# The 10th-step transition matrix
P %^% 10
```

We can now get the 15th-step distribution by running the following code in R:
```{r}
# The 15th-step distribution
mu %*% ( P %^% 15 )
```



## Question 2. 

#### Use the transition matrix above and the initial distribution, $\small{\pi = \left[ \frac{1}{5}, \frac{1}{5}, \frac{1}{5}, \frac{2}{5}\right]}$, to find the 15th-step distribution.



We can find the 15th-step distribution by running the following code in R:
```{r}
# Initial distribution;
pi = c(1/5, 1/5, 1/5, 2/5)

# The 15-th-step distribution
pi %*% ( P %^% 15 )
```



