---
title: "Math 490 HW #7"
author: "Maxwell Levin"
date: "February 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA)
```


## Question 1.

#### Let $\small{X}$ be a discrete r.v. with p.m.f.: $\small{f(1) = \frac{3}{9}, f(2) = \frac{2}{9}, f(3) = \frac{1}{9}, f(4) = \frac{3}{9}}$, and let $\small{Y}$ be a discrete r.v. with p.m.f.: $\small{g(x) = \frac{1}{4}, x = 1,2,3,4}$.

#### a. Find $\small{E(X)}$ and $\small{SD(X)}$.

We calculate $\small{E(X)}$ by

\[
  E(X) = 1\left(\frac{3}{9}\right) + 2\left(\frac{2}{9}\right) + 3\left(\frac{1}{9}\right) + 4\left(\frac{3}{9}\right),
\]
\[
  E(X) = \frac{22}{9} \approx 2.444.
\]


We calculate $\small{Var(X)}$ by

\[
  Var(X) = E(X^2) - E(X)^2,
\]
\[
  E(X^2) = 1^2\left(\frac{3}{9}\right) + 2^2\left(\frac{2}{9}\right) + 3^2\left(\frac{1}{9}\right) + 4^2\left(\frac{3}{9}\right),
\]
\[
  E(X^2) = \frac{68}{9} \approx 7.555,
\]
\[
  Var(X) = \frac{68}{9} - \left(\frac{22}{9}\right)^2 \approx 1.580.
\]

Thus the standard deviation is

\[
  SD(X) = \sqrt{Var(X)} \approx 1.257.
\]


#### b. Use the rejection method to sample from $\small{f(x)}$ using $\small{g(x)}$ as the proposal distribution. Apply the method with 10000 trials. Make a relative frequency table, and report the mean and the standard deviation of the simulated values.

```{r, echo=TRUE}
rejection = function(trials) {
  
  target = c(3/9, 2/9, 1/9, 3/9)
  proposal = c(1/4, 1/4, 1/4, 1/4)
  supremum = 12/9
  results = c()

  for (i in 1:trials) {
    u = runif(1)
    y = sample(1:4,1,prob=proposal)
    if (u <= target[y]/(supremum*proposal[y]) ) {
      results = c(results, y)
    }
  }
  results;
}

trials = 10000
simul = rejection(trials)
table(simul)/length(simul)
```

The mean of our simulation is:
```{r}
mean(simul)
```

The standard deviation is:
```{r}
sd(simul)
```


These are pretty close to what we calculated in part (a).

#### c. For the rejection sampling method used in part (b), what is the probability of acceptance?

We know from class that the probability of acceptance is given by

\[
  Pr\{Accept\} = \frac{1}{M},
\]

where $\small{M}$ is the suprema of $\small{\frac{f(x)}{g(x)}}$. That is,

\[
  M = sup\left(\frac{f(x)}{g(x)}\right) = \frac{12}{9},
\]
\[
  Pr\{Accept\} = \frac{9}{12} = \frac{3}{4}.
\]


We can confirm this experimentally by dividing the length of 'simul' by the number of trials we've run:

```{r, echo=TRUE}

length(simul) / trials

```

This is what we expect, wonderful!


## Question 2.

#### Let random variable $\small{X}$ have the following probability density function:
\[
  f(x) = \left\{
        \begin{array}{ll}
            \frac{x}{50} & \quad 0 \leq x \leq 10; \\
            0 & \quad \text{otherwise.}
        \end{array}
    \right.
\]


#### a. Find $\small{E(X)}$ and $\small{SD(X)}$. 


We calculate $\small{E(X)}$ by

\[
  E(X) = \int_0^{10} \frac{x^2}{50} dx \approx 6.667.
\]

We calculate $\small{Var(X)}$

\[
  Var(X) = E(X^2) - E(X)^2,
\]
\[
  E(X^2) = \int_0^{10} \frac{x^3}{50} dx = 50,
\]
\[
  Var(X) = 50 - (6.667)^2 \approx 5.556.
\]

That is, $\small{SD(X)}$ is

\[
  SD(X) = \sqrt{Var(X)} \approx 2.357.
\]


#### b. Use the rejection method to sample from $\small{f(x)}$ using Uniform(0,10) as the proposal distribution. Apply the method with 10000 trials. Make a histogram, add the density curve to the histogram, and report the mean and the standard deviation of the simulated values.

```{r, echo=TRUE}
target = function(x) {
		ifelse( (x >= 0 & x <= 10), x/50, 0);
}

proposal = function(x) { 0.1; }

rejCont = function(trials) {
  supremum = 5;
	results <- c(); 
  for (i in 1:trials){
	  u<-runif(1);
	  y<-runif(1, 0, 10);
	  if (u <= target(y) / (supremum * proposal(y)) ) {
		  results = c(results, y);
	  }
	}
  results; 
}

trials = 10000
simul = rejCont(trials)
hist(simul, breaks=seq(0, max(simul)+1), prob=T)

x = seq(0, 10, 0.01)
y = ifelse( x >= 0 & x<= 10, x/50, 0 )
lines(x, y, col="blue")
```


The mean is
```{r}
mean(simul)
```

The standard deviation is
```{r}
sd(simul)
```


#### c. For the rejection sampling method used in part (b), what is the probability of acceptance?


The probability of acceptance is $\small{\frac{1}{5}}$ since 5 is the supremum (We find this occurs at $\small{x = 1}$).

We verify this experimentally:
```{r}
length(simul)/trials
```




