---
title: 'Math 490 HW #9'
author: "Maxwell Levin"
date: "February 19, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment=NA)
```

## Question 1. 

#### Consider computing the definite integrate $\small{I = \int_0^\pi sin(x^2)dx}$ using Monte Carlo methods.


#### a. If sample size $\small{N = 4}$, how likely is our Monte Carlo estimate $\small{Z_4^{MC}}$ to yield an estimate of $\small{I}$ within error $\small{\pm 0.01}$? Use R to simulate this probability.

We can use R to experimentally compute this probability by running the following code:
```{r, echo=TRUE}
I = 0.77265
getPr = function(sam, rep) {
  obs = 0
  for (i in 1:rep) {
    x = runif(sam, 0, pi)
    zbar = mean(pi*sin(x^2))
    if (zbar >= I - 0.01 & zbar <= I + 0.01) {
      obs = obs + 1
    }
  }
  obs / rep
}

getPr(4, 10000)
```

Thus we see that we have less than a $\small{1\%}$ chance of achieving an estimate for $\small{I}$ within $\small{\pm 0.01}$. This is terrible!


#### b. If sample size $\small{N = 400}$, how likely is our Monte Carlo estimate $\small{Z_{400}^{MC}}$ to yield an estimate of $\small{I}$ within error $\small{\pm 0.01}$? Use R to simulate this probability.

We can run the following code in R to estimate this probability:

```{r, echo=TRUE}
getPr(400, 10000)
```

We see that we've achieved an $\small{8\%}$ chance of being within two decimal places of our actual result. This is better, but still pretty awful.



#### c. How large should $\small{N}$ be such that our Monte Carlo estimate $\small{Z_{N}^{MC}}$ would yield an estimate of $\small{I}$ within error $\small{\pm 0.01}$ with probability 95%?


We know from class that $\small{N}$ is given by:

\[
  N = \left(\frac{1.96\sigma}{\epsilon} \right)^2,
\]

where $\small{\sigma}$ is our standard deviation, $\small{0.942}$, and $\small{\epsilon}$ is our error bound, i.e., $\small{0.01}$. That is,

\[
  N = \left(\frac{1.96(1.883)}{0.01}\right)^2,
\]
\[
  N \approx 136255.
\]

Thus we see that $\small{N}$ has to be quite large in order to achieve accuracy to two decimal places.










